---
title: "Bellabeat Case Study"
author: "Serena Leung"
date: "7/25/2021"
output: html_document
---

## Ask 

Guiding Questions: 

Goal: Want to gain insight into how non-Bellabeat users use smart devices, then apply insights to one Bellabeat product 


#### What is the problem you are trying to solve?
Why are these users not using Bellabeat?
* Ie: why are they using Fitbit  



#### How can your insights drive business decisions?
* What can we do to get users to start using Bellabeat?  




#### Key Tasks 
1. Business task
+ Why are people using Fitbit instead of Bellabeat?
2. Stakeholders 
+ Bellabeat founders 
+ Strategy, marketing, product team (if insights require them to make changes)  




#### Deliverable
To determine particular trends in Bellabeat’s competitors, we will use the provided Fitbit data (and many include other relevant datasets as well). In this project, we will focus on different factors that competitors have (but Bellabeat don’t have) to ascertain whether this difference has an effect on sales or membership. Ultimately, the goal is to identify trends within our competitor's product and see if we can apply these trends to Bellabeat.



## Prepare

Data Source: 


## Process

Assumptions
* Time to fall asleep corresponds to the same day as the total activity done in the same row 

Note: We will only show correlation


Cleaning Process 

```{r, include=FALSE}
library(tidyverse)
library(lubridate)
library(readr)
library(knitr)
library(kableExtra)

```


```{r, message=FALSE}
# rename df
activity <- read_csv("dailyActivity_merged.csv")
calories <- read_csv("dailyCalories_merged.csv")

# separate date time into data and time columns, drop time column
sleep <- read_csv("sleepDay_merged.csv") %>% 
  separate(SleepDay, c("SleepDay", "other"), sep = " ") %>% select(-other)

```


```{r}
dim(activity)

# Join activity and calories df
act_cal <- left_join(activity, calories, by = c("Id", "ActivityDate" = "ActivityDay", "Calories"))


# Join act_cal df with sleep df, dropped 527 entries with NAs, removed 3 duplicates
fitbit_data <- left_join(act_cal, sleep, by = c("Id", "ActivityDate" = "SleepDay")) %>% 
  na.omit() %>% 
  unique()

dim(fitbit_data)
```


Checking for reasonable ranges and glaringly wrong data points

```{r}
# Sedentary Active Distance should be close to 0 because you're not moving
fitbit_data %>% 
  summarise(min_sedentary = min(SedentaryActiveDistance), max_sedentary = max(SedentaryActiveDistance))
```

```{r}
# Check if all distances add up 
a <- fitbit_data %>% 
  mutate(temp_total = TrackerDistance + LoggedActivitiesDistance + SedentaryActiveDistance + LightActiveDistance + ModeratelyActiveDistance + VeryActiveDistance) %>%
  select(Id, TotalDistance, temp_total)%>% 
  mutate(difference = TotalDistance - temp_total) %>%  # take difference to compare if values are close enough
  filter(difference > 0.1)                             # filter if difference is big

head(a)
```


## Analysis


Basic summary of the data 
```{r}
fitbit_data %>% group_by(Id) %>% 
  summarise(avg_time_fall_asleep = mean(TotalMinutesFallAsleep), 
            avg_steps = mean(TotalSteps), 
            avg_distance = mean(TotalDistance)) %>% 
  arrange(-avg_steps) %>% 
  print(n=10)
```


Compare and contrast clinical data vs Summary data from our dataset


```{r}
clinical <- c("10 - 20", "3000 - 4000", "420 - 540")

fitbit_mean <- lapply(fitbit_data[c("TotalMinutesFallAsleep", "TotalSteps", "TotalMinutesAsleep")], mean)

fitbit_range <- lapply(fitbit_data[c("TotalMinutesFallAsleep", "TotalSteps", "TotalMinutesAsleep")], function(x){
  paste(range(x), collapse = " - ")
})

sum_results <- as.data.frame(rbind(clinical, fitbit_range, fitbit_mean))

colnames(sum_results) <- c("Avg Min to Fall Asleep", "Avg Steps Taken a Day", "Avg Hrs of Sleep")
```


```{r}
sum_results %>% 
  kbl() %>% 
  kable_minimal()
```



#### Goal: Is there a correlation between sleep and exercise?

To better see a possible correlation between sleep and exercise, we can place users into categories depending on their falling asleep rate. 

```{r}
sleep_cat <- cbind(c("Very Slow", "Slow", "Normal", "Fast"), c("< 10 min", "10 - 20 min", "20.001 - 60 min", "> 60.001 min"))

sleep_cat %>% as.data.frame() %>% rename("Sleep Category" = V1, "Range" = V2)

```

Now, let's see a summary of the relevant sleep and exercise data:

```{r}
fitbit_data %>% 
  mutate(fall_asleep_cat = case_when(TotalMinutesFallAsleep < 10 ~ "Fast",
                                                   TotalMinutesFallAsleep >= 10 & TotalMinutesFallAsleep <= 20 ~ "Normal",
                                                   TotalMinutesFallAsleep > 20 & TotalMinutesFallAsleep <= 60~ "Slow", 
                                                   TotalMinutesFallAsleep > 60 ~ "Very Slow") ) %>% 
  group_by(fall_asleep_cat) %>%
  summarise(count = n(), avg_steps = mean(TotalSteps), 
            avg_seden_min = mean(SedentaryMinutes), 
            avg_sleep_dur = mean(TotalMinutesAsleep),
            active_min = mean(VeryActiveMinutes),
            avg_cal_burned = mean(Calories)) 
```




However, it's a little hard to discern trends with a table of numbers, so I built a Tableau dashboard that translates these insights into graphs.  

<iframe src="https://public.tableau.com/views/BellabeatViz/Dashboard1?:showVizHome=no&:embed=true" width="900" height="750"></iframe>




To explore this data vizualization, you can adjust the filters "Fall Asleep" to filter which Fall Asleep Category is shown or "Outlier" to view the data without the outlier in the Very Slow Category.


# fall asleep vs asleep viz is bad!!! 

Conclusion: 

### General Information About the Data

The data is split into 4 categories: Fast, Normal, Slow, and Very Slow. The Fast and Very Slow make up about 25% of the data, as shown in the Grouped by ID graph in "Count of Fall Asleep Cat". The Slow category makes up the most data points, which is reasonable since we expect most our sample to fall between the extreme ranges. 


### Slow Category 

Data 

### Fast Category 



FAST: People who fall asleep more quickly, sleep less than 5 hrs, and have more sedentary minutes 
### maybe don't feel like exercising bc too tired, fall asleep quickly bc already sleep deprived, actually burn the most calories 
### avg_seden_min higher mebbe bc they're awake for more hours of the day

VERY SLOW: burn the least calories, spend the least time working out

- outlier 


